# Data sources


## Time Series Data Library
Rob J. Hyndman maintained the Time Series Data Library for a long time. Now you can get the data from the R package tsdl. It used to be available fron [DataMarket](https://datamarket.com/data/list/?q=provider:tsdl) but the site has been acquired by Qlik.  The collection consists of about 560 time series of all various variables. Unfortunately, the tsdl package is not on CRAN but can be installed using the following R commands:

```{r eval=FALSE} 
install.packages("devtools")
library(devtools)
install_github("FinYang/tsdl")
```

## DataMarket
DataMarket was a very good source for data. Unfortunately it has been acquired by Qlik and the informaiton about how to access the data nowadays is lacking. They do have an [acadmic program](https://www.qlik.com/us/academicprogram) but it is not clear what comes with it.

## Quandl

The Quandl ata source has many economic and financial datasets. For some data you need to have a premoium account but for many you do not need to. You can find the series here: [Quandl](https://www.quandl.com/search) where you can browse for the data you want. Once you found what you looking for, press Expand and then there is a R button on the right hand side and clicking on it yields the R code to download it.

To install the Quandl package:

```{r eval=FALSE} 
install.packages("Quandl")
library(Quandl)
```

For example, to download the Big Mac index for Sweden

```{r eval=FALSE} 
Quandl("ECONOMIST/BIGMAC_SWE")
```

More information can be found at [Quandl R Guide](https://docs.quandl.com/docs/r).

## Dataseries
[dataseries.org](http://www.dataseries.org/) is a comprehensive data archive for Switzerland.


```{r eval=FALSE} 
install.packages("dataseries")
library(dataseries)
```

For ecah series you get the R code. For example CPI

```{r eval=FALSE} 
dataseries::ds("CPI.LD2010100")
```

## pdfetch

Abiel Reinhart has made an R package to make it easier to download public economic and financial data.  The sources are: St Louis Fed's FRED system, Yahoo Finance, the US Bureau of Labor Statistics, the US Energy Information Administration, the World Bank, Eurostat,the European Central Bank, the Bank of England, the UK's Ofﬁce of National Statistics, Deutsche Bundesbank, and INSEE.

It is availalbe at CRAN and can be installed using:

```{r eval=FALSE} 
install.packages("pdfetch")
library(pdfetch)
```

Depending on the source differnt functions needs to be used and the identifiers need to be found att the various sources.

## IPUMS

IPUMs provides data from censuses and surveys from around the world. There are huge amounst of data, e.g. IPUMS International has data for almost 100 countries. To use IPUMS you need to register but the data is free of charge. It can be found here: [IPUMS](https://www.ipums.org/). There is, to my knowledge, no nice R interface.

## ESS

European Social Surveys is a biennial cross-country survey of attitudes and behaviour. The first round was in 2001 and it is now at round 8. Hence, it is possible to both analyze changes over time as well as differences between countries. The access is free of charge but a registration is nessisary. 

## Penn World Table

The Penn World Table consists of data on relative income and other econmic indicators (almost 30) for more than 180 countries between 1950 and 2014. It is developed by Robert Summers and Alan Heston (amongst others) for the purpose of comparing national accounts. It is administrated by Groningen Growth and Development Centre, Faculty of Economics and Business, University of Groningen. There is a R package pwt9 (available at CRAN) which facilitates access to the data.


## kaggle
kaggle is one of the largest collections of datasets with more than 15,000 datasets.It consists of all varoius types of data and you do not need to register to access them. 
[kaggle](https://www.kaggle.com/datasets)

## Forecasting competitions
The Makridakis competitions or the M-competetitions are a series of forecasting competitions with the aim of evaluating forecasting methods. The first one was in 1982 while the fourth 2018. The number of series have varied as well as the seasonal frequency. The first consisted of 1001 series and the fourth 100,000, while the second only 29. The third had 3003 series. 
The data from the first nd third competitions may be found in the R package Mcomp. 

The organisors of the third M-competition tried to get researcher in artificial neural networks to participate but due to the many series and the computational burdon of ANN only one did and performed poorly. This inspuired the NN3-competition which was based on 111 of the M3-competition data (same source but shifted in time). These data can be found at kaggle.

Yet another forcasting competition was the 2010 Tourist competition. As they couldn't get permission for all data explanatory variables are not included in the public data. The data can be found at kaggle or through the R-package Tcomp.

## Hyndman data
Rob J Hyndman is a prominent profesor at Monash University doing research in forcasting. He has authored many books and data can be found at the following R-packages: fma, expsmooth, fpp, fpp2. 

## UC Irvine Machine Learning Repository

A nice collection of 468 data sets can be found at [UC Irvine Machine Learning Repository](https://archive.ics.uci.edu/ml/index.php). The data sets are not recent but there is a large variaty. There are data sets on omages, census data, el Nino etc, suitably for data mining exercises.

## yelp

A source of a very specific data set is the [yelp data](https://www.yelp.com/dataset). It contains of almost 9 gigabytes of uncompressed data (3.6 comrpessed) on reviews etc on various services, like restaurants etc. The photos can also be downloaded. The data is free for academic purposes. The data can be access using JSON which is a lightweight data-interchange format. R can read JSON files using the packages rjson, RJSONIO and jsonlite. 

## Official US data

US has made availible a huge number of data sets, about 240,000, through [data.gov](https://www.data.gov). The data are on differnt levels such as cities, state and federal on various topics. it can be everything from estimated  average prices for over 150 commonly consumed fresh and processed fruits and vegetables to crime rates for Chicago to the present. The Chicago crime rates data consists of all crimes in the Chicago area including information on type of crime, case id, location  etc on each crime committed. The data can be downloaded in csv, rdf, json or xml formats.

Also [US Census Bureau ](https://www.census.gov/data.html) is a good source of data. 

## 538

The site [FiveThirtyEight](https://data.fivethirtyeight.com/) is a site focusing on opinion polls, politics, economics and sports. It originated as a polling aggregation website created by Nate Silver. Compared to many other sites 538 share data and code behind some of their articles and graphics. They have data which is not common such as NBA predictions, "How popular is Trump" and Senate forecasts.

## Football data

A main source of football data is [football-data.co.uk](http://www.football-data.co.uk/data.php) where you can find historical data on all main leagues in the world and some minor leagues. E.g. for England you can getmatch results for the four highest leagues from season 93/94. The newer data includes much more data such as shots, shots on target, woodwork hits, fouls committed besides half and full time results. The data is in csv format and needs to be downloaded league by league, season by season. Some of the data can be accessed using the GitHub page [sports-data](https://datahub.io/sports-data) but it seems to be restricted to about the last 10 seasons.

The site [football-data.org](https://www.football-data.org/) offers access using e.g. R code but requires registration for 12 leagues. Monthly fee yileds access to up to 144 competitions. There is no information on the site on when the historical data starts.

A very nice R package is [engsoccerdata](https://github.com/jalapic/engsoccerdata) with e.g. data on the four highest leagues in England from 1888- 2017. The package is not on CRAN and the following code may be used:

```{r eval=FALSE} 
library(devtools)
install_github('jalapic/engsoccerdata', username = "jalapic")
library(engsoccerdata)

data(package="engsoccerdata")    # lists datasets currently available
```


## Machine learning data set respository

More than 800 data sets can be found at [mldata.org](http://mldata.org/repository/data/). There are a variety of topics, e.g. Plant classification, Book Evaluation, andLinux kernel statistics.  The formats are not coherent amongst the data sets but common are XML, csv and mat. A related site is [mloss.org](https://mloss.org/software/) with machine learning code.

## Open ML

[Open ML](https://www.openml.org/search?type=data&q=tags.tag:OpenML100) has about 90 data set suitable for machine learning. There are data sets on images (e.g. handwritten numerals, the Breast Cancer Wisconsin (Original) Data Set), speach recongnition (e.g. letters, nasal vs oral sound), and other interesting data sets such as the widely used Australian New South Wales Electricity Market data set. There is an OpenML R package dowloadable from CRAN. Instructions can be found [here](https://openml.github.io/OpenML/R-guide/) including a public read-only API key.

## Academic Torrent

The other sites host the data while the [Academic Torrents](http://academictorrents.com/browse.php?cat=6) use the BitTorrent protocol for peer to peer file sharing. This is a method more famously used to share movies. Here, the use of the method is legal and implies high speed downloading of the data sets. Ther are more than 42 Terabytes of all kind of data. There is an old R package [atdownload](https://github.com/liuyifei0226/atproject) as well as old [java tool](https://josephpcohen.com/w/academic-torrent-download-tool-atdown/) but it is inclear if they are still working.

## Swedish National Data Service

[SND](https://snd.gu.se/en/catalogue/search) is a public data base fort Swedish research data. Some data is hosted at the SND site while other is accessed from external actors. Also the level of access differs, some data you can just download, sometiomes you ned toregister first to be able to donwload. For some data sets you need to ask for permission from the data owner. 

## EU Data

The [EU Open Data Portal](http://data.europa.eu/euodp/en/data/) has as purpose "EU institutions are invited to make their data publicly available whenever possible". The data can be used free of charge and without copyright restrictions as long as you acknowledge the source. There is no R package but curl can be used to access the [API](http://data.europa.eu/euodp/en/developerscorner). The data origins from EU authorities, hence they are related to them. Examples include Gender Equlity Index, Pupils and students enrolled by education level, sex and age, and Nights spent at tourist accommodation establishments.

Compared to the EU Open Data Portal, the [European Data Portal](https://www.europeandataportal.eu/data/#/datasets) consists of data from the indvidual EU member countries. 